Hierarchically Supervised Latent Dirichlet Allocation

### 2011-06-02 paper/experiments.tex
fixed fig
edits to experiments

We experimented with HSLDA for prediction in two domains: predicting medical
diagnosis codes from hospital discharge summaries and predicting product
categories from Amazon.com product descriptions.

@@

Discharge summaries are authored by clinicians to summarize the course of a
hospitalized patient. The summaries typically contain a record of the patient
complains, findings and diagnoses, along with treatment and hospital course.
For each admission trained medical coders review the information in the
discharge summary and assign a series of diagnoses codes. Coding follows the
ICD-9-CM controlled terminology, an international diagnostic classification for
epidemiological, health management, and clinical
purposes.\footnote{http://www.cdc.gov/nchs/icd/icd9cm.htm} As such, the ICD-9
codes constitute a labeling of a patient's diagnoses based on a discharge
summary. The ICD-9 codes are organized in a rooted-tree structure, with each
edge representing an is-a relationship between parent and child, such that the
parent diagnosis subsumes the child diagnosis. For example, the code for
{}``Pneumonia due to adenovirus'' is a child of the code for {}``Viral
pneumonia,'' where the former is a type of the latter.  It is worth noting that
the coding can be noisy. Human coders sometimes disagree~\cite{Challenge07},
tend to be more specific than sensitive in their
assignments~\cite{Birmetal2005}, and sometimes make
mistakes~\cite{FarzandipourEtAl10}.

The task of automatic ICD-9 coding has been investigated in the clinical
domain. Much of the work was triggered by the 2007 medical NLP community
challenge~\cite{Challenge07}. The data in the challenge, however, differs from 
ours in its scope. The datasets were smaller (1,000 training and 1,000 testing
documents) and focused on radiology reports with a restricted number of ICD-9
codes (45 of them, compared to 7K+ in our dataset). Methods ranged from manual
rules to online learning~\cite{Crammer2007,Goldstein2007,Farkas2008}.
Other work had leveraged larger datasets and experimented with K-nearest
neighbor, Naive Bayes, support vector machines, Bayesian Ridge Regression, as
well as simple keyword mappings, all with promising
results~\cite{LarkeyCroft95,RibeiroNeto2001,PakhomovEtAl06,Ruch2008,Lita2008}.

Our dataset was gathered from the clinical data warehouse of
NewYork-Presbyterian Hospital. It consists of 6,000 discharge summaries and
their associated ICD-9 codes (7,298 distinct codes overall), representing all
the discharges from the hospital in 2009. Summaries have 8.39 associated ICD-9
codes on average (std dev=5.01) and contain an average of 536.57 terms after
preprocessing (std dev=300.29). We split our dataset into 5,000 discharge
summaries for training and 1,000 for testing.

The text of the discharge summaries was tokenized with
NLTK.\footnote{http://www.nltk.org} Vocabulary was determined as the top 10,000
tokens with highest document frequency (exclusive of names,
places and other identifying numbers). Each discharge summary is thus
represented as counts over the 10,000-word vocabulary. The study was approved
by the Institutional Review Board and follows HIPAA (Health
Insurance Portability and Accountability Act) privacy guidelines.

@@

Amazon.com, an online retail store, organizes its catalog of 
products in a hierarchy and provides product descriptions
for most products in their catolog.  Products can be discovered by users
through query-based searching or through product category exploration. Top-level
product categories are displayed on the front page of the website and lower
level categories can be discovered by choosing one of the top-level categories.
Products can exist in multiple locations in the hierarchy.

@@

We were able to deduce the structure of the
hierarchy for the Amazon.com products because all ancestors in the
hierarchy were included with each category label. For example, {}``DVD / Genres
/ Science Fiction \& Fantasy / Classic Sci-Fi'' is a single product category 
for the DVD, {}``The Time Machine.''

@@

We evaluated HSLDA along with three other closely related models against 
these two datasets. The comparison models included sLDA with independent 
regressors (hierarchical constraints on labels ignored), HSLDA fit by 
performing LDA followed by tree-conditional regressions, and HSLDA fit 
with fixed random regression parameters. These models were chosen to 
highlight several aspects of the model including performance in the absence 
of hierarchical constraints, the effect of the combined inference procedure, and regression
performance attributable solely to the hierarchical constraints.

sLDA with independent regressors is the most salient comparison model
for our work. The distinguishing factor between HSLDA and sLDA is the
addition structure imposed on the label space, a distinction that we
hypothesize will result in a difference in predictive performance. 
Aside from this, all other features of sLDA are preserved between the two models.

@@

The last comparison model is HSLDA with fixed and randomly selected regression
parameters. There is a baseline benefit that the structure in the label space 
provides for the prediction of labels. This comparison model is intended to
quantify the contribution of the structure alone.

@@

To evaluate the performance of these models, we established a gold standard for comparison.
For each dataset, a held out set of 1000 documents and labels were
reserved for evaluation and predictive performance was evaluated against a standard derived
from the observed labeling. Specifically, ancestors of observed nodes were ignored, observed nodes 
were considered positive and unobserved nodes were considered to be negative. This method of 
defining positive and negative labels was chosen to be as fair as possible to all models
being compared.  In particular, since the sLDA model does not enforce the hierarchical
constraints, the models can be compared on a more equal footing by considering only on the observed labels
as being positive despite the fact that ancestors must also be positive. The gold standard 
defined in this way will likely lead to a slight overestimation of the number of
false positives. It is known that ICD-9 codes lack sensitivity and their use as a
gold standard could lead to correctly positive predictions being labeled as false positives.
However, given that the label space is often large (as in our examples) it is a moderate assumption that
erroneous false positives should not skew results significantly.

@@

Figure \ref{fig:main_results} shows the predictive performance of HSLDA 
relative to the three comparison models as a function of the prior mean
on regression coefficient. Figure \ref{fig:1a} demo.

@@

\begin{figure}[t]
%tbp] %  figure placement: here, top, bottom, or page
 \centering \includegraphics[scale=0.4]{figs/ROC_comparison_leafs} \caption{ROC Curve for clinical data}

>>>>

We applied HSLDA to data from two domains: predicting medical
diagnosis codes from hospital discharge summaries and predicting product
categories from Amazon.com product descriptions.

@@

Discharge summaries are authored by clinicians to summarize patient hospitalization course. 
The summaries typically contain a record of patient
complaints, findings and diagnoses, along with treatment and hospital course.
For each hospitalization, trained medical coders review the information in the
discharge summary and assign a series of diagnoses codes. Coding follows the
ICD-9-CM controlled terminology, an international diagnostic classification for
epidemiological, health management, and clinical
purposes.\footnote{http://www.cdc.gov/nchs/icd/icd9cm.htm}  The ICD-9 codes are organized in a rooted-tree structure, with each
edge representing an is-a relationship between parent and child, such that the
parent diagnosis subsumes the child diagnosis. For example, the code for
{}``Pneumonia due to adenovirus'' is a child of the code for {}``Viral
pneumonia,'' where the former is a type of the latter.  It is worth noting that
the coding can be noisy. Human coders sometimes disagree~\cite{Challenge07},
tend to be more specific than sensitive in their
assignments~\cite{Birmetal2005}, and sometimes make
mistakes~\cite{FarzandipourEtAl10}.

The task of automatic ICD-9 coding has been investigated in the clinical
domain.  Methods range from manual
rules to online learning~\cite{Crammer2007,Goldstein2007,Farkas2008}.
Other work had leveraged larger datasets and experimented with K-nearest
neighbor, Naive Bayes, support vector machines, Bayesian Ridge Regression, as
well as simple keyword mappings, all with promising
results~\cite{LarkeyCroft95,RibeiroNeto2001,PakhomovEtAl06,Ruch2008,Lita2008}.

%Much of the work was triggered by the 2007 medical NLP community
%challenge~\cite{Challenge07}.
%The data in the challenge, however, differs from 
%ours in its scope. The datasets were smaller (1,000 training and 1,000 testing
%documents) and focused on radiology reports with a restricted number of ICD-9
%codes (45 of them, compared to 7K+ in our dataset).

Our dataset was gathered from the clinical data warehouse of
a large metropolitan hospital. It consists of 6,000 discharge summaries and
their associated ICD-9 codes (7,298 distinct codes overall), representing all
the discharges from the hospital in 2009. Summaries have 8.39 associated ICD-9
codes on average (std dev=5.01) and contain an average of 536.57 terms after
preprocessing (std dev=300.29). We split our dataset into 5,000 discharge
summaries for training and 1,000 for testing.

The text of the discharge summaries was tokenized with
NLTK.\footnote{http://www.nltk.org} A fixed vocabulary was formed by taking
the top 10,000 tokens with highest document frequency (exclusive of names,
places and other identifying numbers). The study was approved
by the Institutional Review Board and follows HIPAA (Health
Insurance Portability and Accountability Act) privacy guidelines.

@@

Amazon.com, an online retail store, organizes its catalog of 
products in a mulitply-rooted hierarchy and provides textual product descriptions
for most products.  Products can be discovered by users
through query-based searching or through product category exploration. Top-level
product categories are displayed on the front page of the website and lower
level categories can be discovered by choosing one of the top-level categories.
Products can exist in multiple locations in the hierarchy.

@@

%We were able to deduce the structure of the
%hierarchy for the Amazon.com products because all ancestors in the
%hierarchy were included with each category label. For example, {}``DVD / Genres
%/ Science Fiction \& Fantasy / Classic Sci-Fi'' is a single product category 
%for the DVD, {}``The Time Machine.''

@@

We evaluated HSLDA along with three other closely related models against 
these two datasets. The comparison models included sLDA with independent 
regressors (hierarchical constraints on labels ignored), HSLDA fit by first
performing LDA then fitting tree-conditional regressions. These models were chosen to 
highlight several aspects of HSLDA including performance in the absence 
of hierarchical constraints, the effect of the combined inference, and regression
performance attributable solely to the hierarchical constraints.

sLDA with independent regressors is the most salient comparison model
for our work. The distinguishing factor between HSLDA and sLDA is the
addition structure imposed on the label space, a distinction that we
hypothesized would result in a difference in predictive performance.

@@

%The last comparison model is HSLDA with fixed and randomly selected regression
%parameters. There is a baseline benefit that the structure in the label space 
%provides for the prediction of labels. This comparison model is intended to
%quantify the contribution of the structure alone.

@@

To evaluate the performance of these models, we established a gold standard for comparison.
For each dataset, a held out set of 1000 documents and labels were
reserved for evaluation and predictive performance was evaluated against a standard derived
from the observed labeling. To make the comparison as fair as possible, 
ancestors of observed nodes were ignored, observed nodes 
were considered positive and descendents of observed nodes were considered to be negative. This method of 
defining positive and negative labels was chosen to be as fair as possible to all models
being compared.  In particular, since the sLDA model does not enforce the hierarchical
constraints, the models can be compared on a more equal footing by considering only on the observed labels
as being positive despite the fact that ancestors must also be positive. The gold standard 
defined in this way will likely lead to a slight overestimation of the number of
false positives. It is known that ICD-9 codes lack sensitivity and their use as a
gold standard could lead to correctly positive predictions being labeled as false positives.
However, given that the label space is often large (as in our examples) it is a moderate assumption that
erroneous false positives should not skew results significantly.

@@

Figure \ref{fig:main_results} shows the predictive performance of HSLDA 
relative to the three comparison models as a function of the prior mean
on regression coefficients as a receiver operating characteristic (ROC) curve. 
For low values of the auxiliary variable
threshold, the models predict labels in a more sensitive and less specific
manner, creating the points in the upper right corner of the ROC curve. As
the auxiliary variable threshold is increased, the models predict in a less
sensitive and more specific manner, creating the points in the lower left
hand corner of the ROC curve. For all values of the prior mean in both datasets
HSLDA outperforms sLDA with independent regressors. In the case of HSLDA with
separately trained regression, HSLDA outperforms in the clinical dataset but
performs equally well across the board with the retail product dataset.

@@

\begin{figure}[t]
%tbp] %  figure placement: here, top, bottom, or page
 \centering \includegraphics[scale=0.4]{figs/ROC_comparison_leafs} \caption{ROC Curve for clinical data}
\label{fig:clinical_roc} 
\end{figure}
